---
title: "Continuous data example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{continuous_data_example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(foocafeReliability)
library(rstan)
```


## Build model  (based on example 4.3 from Bayesian Reliability)

Failure time data is commonly used to assess component reliability. This is a
record of the continuous time until a component failed. In general failure time
data record "time to some event", with other examples including "time to death"
in survival analysis, and "time to interrupt" in software reliability.

There are several standard failure time models and here we consider the
exponential, Weibull, and lognormal failure time models to model LCD projector
lamp failure times.

In business and educational settings, computer presentations use liquid
crystal display (LCD) projectors. The most common failure mode of these
projectors is the failure of the lamp. Many manufacturers include the
"expected" lamp life in their technical specification documents, and one
manufacturer claims that users can expect 1,500 hours of projection time
from each lamp used under "normal operating conditions". To test this claim,
a large private university placed identical lamps in three projector models
for a total of 31 projectors. The university staff recorded the number of
projection hours (as measured by the projector) when each lamp burned out.
This dataset are the results from that experiment.

Exponential model

Assuming a constant failure rate, we first consider modelling the lamp failure
times by an exponential distribution with rate parameter $\lambda$ and use the 
conjugate prior distribution for $\lambda$,
$$
\lambda \sim Gamma(\alpha, \beta)
$$
One way to choose the values for $\alpha$ and $\beta$ is to use the
manufacturers best guess for the Mean Time To Failure, or MTTF $(1/\lambda)$
and use a large standard deviation around this specification. We take the
manufacturers statement of "about 1500 hours" of lamp life as a prior mean
lamp life of 1500 hours, and interpret "about" as not being very certain. Using
moment matching (see Hamada et al. p93 for details), we obtain suitable values
for alpha and beta; $\alpha \approx 2.5$ and $\beta \approx 2350$.

We now build the model using the rstan package. Lambda is the key parameter in
this model. MTTF is simply a transformation of lambda. The variable
'lambda_prior' is defined in order to the posterior for lambda later on.

```{R}

modelString <- "
  data {
    int<lower=0> Nobs;
    real<lower=0> projection_hours[Nobs];
  }

  parameters {
    real<lower=0> lambda;
    
    real<lower=0> lambda_prior;
  }
  
  transformed parameters {
    real MTTF;
    
    MTTF = 1 / lambda;
  }

  model {
    projection_hours ~ exponential(lambda);

    lambda ~ gamma(2.5, 2350);
    
    lambda_prior ~ gamma(2.5, 2350);
  }
"

lcd_exponential_model <- rstan::stan_model(model_code = modelString)

```

Then we fit the model to the LCD failures data.
```{R}

lcd_fail_list <- list(Nobs = NROW(lcd_projector_failures$projection_hours),
                          projection_hours = lcd_projector_failures$projection_hours
                          )

lcd_exponential_output <- rstan::sampling(lcd_exponential_model, data = lcd_fail_list, chains = 4, iter = 4000, control = list(adapt_delta = 0.9))

```

The posterior distributions can be displayed using stan_hist
```{R}

stan_hist(lcd_exponential_output, c("lambda", "lambda_prior", "MTTF"))

```
Comparing the prior (dashed line) with the posterior (solid line)

```{R}

lcd_exponential_output_vals <- extract(lcd_exponential_output)

data <- data.frame(lcd_exponential_output_vals["lambda"],
                   lcd_exponential_output_vals["lambda_prior"],
                   lcd_exponential_output_vals["MTTF"])

ggplot(data = data) +
  geom_density(aes(lambda)) +
  geom_density(aes(lambda_prior), linetype = "dashed")
```



```{R}
ggplot(data = data) +
  geom_density(aes(MTTF))

```


```{R}

t = seq(1,3500,100)
lambda_vals <- lcd_exponential_output_vals[["lambda"]]

reliability <- data.frame(matrix(0,
                          nrow = NROW(t),
                          ncol = NROW(lambda_vals)))

for (ii in 1:NROW(reliability)) {
  print(ii)
  reliability[ii,] = exp(-lambda_vals * t[ii])
}

ggplot(data = reliability) +
  geom_density(aes(lambda)) +
  geom_density(aes(lambda_prior), linetype = "dashed")


```


```{R}

costTableList <- list()

for (ii in 1:NROW(lcd_exponential_output_vals$lambda)){

  thisLambda <- lcd_exponential_output_vals$lambda[ii]
  
  get_reliability <- function(thisLambda){
    f1 <- function(t){
      exp(-thisLambda * t)
    }
    return(f1)
  }

  reliability <- get_reliability(thisLambda)

  costTable <- data.frame("t" = t,
                          "reliability" = numeric(NROW(t)))

  costTable$reliability <- reliability(t)
  
  costTableList[[ii]] <- costTable
}

temp <- lapply(costTableList, '[', 2) %>% dplyr::bind_cols()

data <- data.frame(t = t,
                   meanVal = apply(temp, 1, mean),
                   LQ = apply(temp, 1, function(x) {quantile(x, 0.1)}),
                   UQ = apply(temp, 1, function(x) {quantile(x, 0.90)}))


ggplot(data = data, aes(t)) +
  geom_line(aes(y = meanVal)) +
  geom_ribbon(aes(ymin = LQ, ymax = UQ), alpha = 0.05)
  

```



## Evaluate goodness of fit

```{R}
params <- data.frame(rstan::extract(lcd_exponential_output)["lambda"])
colnames(params) <- "rate"

t <- bayesian_chi_squared_test(y = lcd_projector_failures$projection_hours,
                               distribution_fun = pexp,
                               params = params,
                               data_type = "continuous")

```

### Weibull fit:

## Build model
```{R}

modelString <- "
  data {
    int<lower=0> Nobs;
    vector<lower=0>[Nobs] projection_hours;
  }

  parameters {
    real<lower=0> shape;
    real<lower=0> scale;
    
  }

  model {
    target += weibull_lpdf(projection_hours | shape, scale);

    shape ~ gamma(1, 1);
    scale ~ gamma(2.5, 0.006);
  }
"

lcd_weibull_model <- rstan::stan_model(model_code = modelString)

lcd_fail_list <- list(Nobs = NROW(lcd_projector_failures$projection_hours),
                          projection_hours = lcd_projector_failures$projection_hours
                          )

lcd_weibull_output <- rstan::sampling(lcd_weibull_model, data = lcd_fail_list, chains = 4, iter = 4000, control = list(adapt_delta = 0.9))

```

## Evaluate goodness of fit
```{R}

n <- NROW(lcd_projector_failures)
K <- round(n^0.4)
a <- seq(0,1,1/K)
p <- diff(a)
m <- numeric(K)
output_vals <- rstan::extract(lcd_weibull_output)

chi_val <- qchisq(0.95,K-1)
r_b <- numeric(NROW(output_vals$shape))

for (jj in 1:NROW(r_b)) {
  
  thisShape <- output_vals$shape[jj]
  thisScale <- output_vals$scale[jj]
  
  probabilities <- pweibull(lcd_projector_failures$projection_hours, thisShape, thisScale)
  
  for (ii in 1:length(m)){
    m[ii] <- sum(probabilities > a[ii] & probabilities < a[ii+1])
  }
  
  r_b[jj] <- sum(((m - n*p)^2)/(n*p))

}

(sum(r_b > chi_val)/NROW(r_b) * 100)

```


### lognormal fit:

## Build model
```{R}

modelString <- "
  data {
    int<lower=0> Nobs;
    vector<lower=0>[Nobs] projection_hours;
  }

  parameters {
    real mu;
    real sigma_2;
  }
  
  transformed parameters {
    real<lower=0> sigma;
    
    sigma = sqrt(sigma_2);
  }

  model {
    
    target += lognormal_lpdf(projection_hours | mu, sigma);

    mu ~ normal(6, 5);
    sigma_2 ~ inv_gamma(6.5, 23.5);
  }
"

lcd_lognormal_model <- rstan::stan_model(model_code = modelString)

lcd_fail_list <- list(Nobs = NROW(lcd_projector_failures$projection_hours),
                          projection_hours = lcd_projector_failures$projection_hours
                          )

lcd_lognormal_output <- rstan::sampling(lcd_lognormal_model, data = lcd_fail_list, chains = 4, iter = 4000, control = list(adapt_delta = 0.9))

```


## Evaluate goodness of fit
```{R}

n <- NROW(lcd_projector_failures)
K <- round(n^0.4)
a <- seq(0,1,1/K)
p <- diff(a)
m <- numeric(K)
output_vals <- rstan::extract(lcd_lognormal_output)

chi_val <- qchisq(0.95,K-1)
r_b <- numeric(NROW(output_vals$mu))

for (jj in 1:NROW(r_b)) {
  
  thisMu <- output_vals$mu[jj]
  thisSigma <- output_vals$sigma[jj]
  
  probabilities <- plnorm(lcd_projector_failures$projection_hours, thisMu, thisSigma)
  
  for (ii in 1:length(m)){
    m[ii] <- sum(probabilities > a[ii] & probabilities < a[ii+1])
  }
  
  r_b[jj] <- sum(((m - n*p)^2)/(n*p))

}

(sum(r_b > chi_val)/NROW(r_b) * 100)

```
